Highly Available Transactions: Virtues and Limitations

### 2013-03-31 writeup/intro.tex
new intro

Our main contributions are as follows. We fill gap between, on one
hand, the plethora of previously proposed isolation and consistency
models, and, on the other hand, the goal of high availability.  We
classify which, among the wide array models, are highly available,
denoting them as {\em Highly Available Transactions (HATs)}. In doing
so, we demonstrate that many \textit{implementations} of HAT semantics
are not highly available but that their unavailability is not
\textit{inherently} required. Our investigation shows, somewhat
surprisingly, that besides serializability, only Snapshot Isolation
(SI) and Repeatable Reads (RR) are unavailable. We also show that many
of the weaker data consistency models from distributed systems are
HAT-compliant.

More precisely, highly available systems cannot detect conflicts
between concurrent updates (as needed for preventing Lost Updates or
Write Skew phenomena). However, Read Committed isolation,
Transactional Atomicity, and many of the weaker data consistency
models from database and distributed systems can be implemented in a
highly available manner. For several of these guarantees, such as
Transactional Causal Predicate Cut Isolation, we consider a modified
form of high availiabity in which clients ``stick to'' (i.e., have
affinity with) at least one server (a property which is often implicit
in the distributed systems literature~\cite{herlihy-art, cops, eiger,
  cac} but which requires \textit{explicit} consideration in a
client-server replicated database context). This sticky availability
is easily implementatable and widely employed~\cite{cops, eiger,
  vogels-defs} but is strictly less available than traditional
definitions of high availability.

>>>>

Our main contributions are as follows. We fill gap between, on one
hand, the plethora of previously proposed isolation and consistency
models, and, on the other hand, the goal of high availability.  We
classify which, among the wide array models, are highly available,
denoting them as {\em Highly Available Transactions (HATs)}. In doing
so, we demonstrate that many \textit{implementations} of HAT semantics
are not highly available but that their unavailability is not
\textit{inherently} required. Our investigation shows, somewhat
surprisingly, that besides serializability, only Snapshot Isolation
(SI) and Repeatable Reads (RR) are unavailable, with most other
isolation levels being available. We also show that many of the weaker
data consistency models from distributed systems are HAT-compliant.

More precisely, the reason why fairly strong models like Snapshot
Isolation (SI) and Repeatable Reads (RR) are not highly available is
that they require detecting conflicts between concurrent updates (as
needed for preventing Lost Updates or Write Skew phenomena). However,
Read Committed isolation, Transactional Atomicity, and many of the
weaker data consistency models from database and distributed systems
can be implemented in a highly available manner. For several of these
guarantees, such as Transactional Causal Predicate Cut Isolation, we
consider a modified form of high availiabity in which clients ``stick
to'' (i.e., have affinity with) at least one server (a property which
is often implicit in the distributed systems
literature~\cite{herlihy-art, cops, eiger, cac} but which requires
\textit{explicit} consideration in a client-server replicated database
context). This sticky availability is easily implementatable and
widely employed~\cite{cops, eiger, vogels-defs} but is strictly less
available than traditional definitions of high availability.
